## DS 785: Capstone Project

## Daniel Clements
## University of Wisconsin-Eau Claire

## Presented to Dr. Tracy Bibelnieks, University of Wisconsin-Superior.

## This code uses uses a two-way ANOVA test to evaluate if there are significant differences in responses to
## three questions posed in an IA2030MX survey on artificial intelligence in Mexico. 

---

# Load libraries.

```{r}
library( dplyr )
library( AICcmodavg )
library( ordinal )

```


## Import data files.

```{r}
# Set the working directory.
setwd( "D:/##Backup/Academics/UW-EC/DS Courses/785/Capstone/Data/#R/IA2030MX" ) 

# Import information and data for Survey Item 6-1.
Item_6_1 <- read.csv( "IA2030MX_Q6-1_DC.csv" )
Item_6_2 <- read.csv( "IA2030MX_Q6-2_DC.csv" )
Item_6_3 <- read.csv( "IA2030MX_Q6-3_DC.csv" )
Item_6_ANOVA <- read.csv( "IA2030MX_Q6_DC_ANOVA.csv" )



```

## Create a dataframe including only the sentiment scores of each Item.

```{r}
# 'Item_6_1' has the greatest length. Add 'NA' values to the absent rows of 'Item_6_2' and 'Item_6_3'.
max_ln <- length( Item_6_1$Sent.Scale )

# Combine the 'Item_6_1', 'Item_6_2', and 'Item_6_3' Sentiment Score rows.
ss_6_df <- data.frame( I6_1 = c( Item_6_1$Sent.Scale ),
                       I6_2 = c( Item_6_2$Sent.Scale, rep( NA, max_ln - length( Item_6_2$Sent.Scale ) ) ),
                       I6_3 = c( Item_6_3$Sent.Scale, rep( NA, max_ln - length( Item_6_3$Sent.Scale ) ) ) )

# Cast 'ss_6_df' as a matrix.
ss_6_df_mat <- as.matrix( ss_6_df )

# Test whether atomic vectors exist in 'ss_6_df_mat'.
is_vector( ss_6_df_mat )

```

## Perform two-way ANOVA testing.

```{r}
# Perform two-way ANOVA testing.
ss_6_anal_of_var <- aov( ï..Q6_1 ~ Q6_2 + Q6_3, data = Item_6_ANOVA )

print( Item_6_ANOVA )

```

## Display the results.

```{r}
# Display two-way ANOVA results, Q6_1 versus Q6_2 and Q6_3.
summary( ss_6_anal_of_var )
```

